Investigating the Effect of Dataset Representivity for Image Colourisation

### 2019-10-12 Paper/paper.tex
Cleaned directory
Added related works

\documentclass[conference]{IEEEtran}
% \IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\graphicspath{{./images/}}
\usepackage{textcomp}
\usepackage{subfig}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

@@

%Explain notation and setup for regression/classification task in general i.e. objective function

%Describe LAB; classes, etc.

%Describe eval criteria

%Describe neural network, backprop, Insert image if space
\subsection{Colour space}

To make our lives easier, instead of using the typical RGB color space, we use the CIELAB color space. This makes it convenient when seperating the gray and color channels due to the fact that the 'L' channel in LAB is simply the grayscale image. We can then pass this grayscale input (1 x W x H) through our model and compare the output (2 x W x H) to the original 'AB' channels of the image. These 'AB' channels make up the color and simply show the red-green and blue-yellow shift respectively. HSB would be another option when choosing a suitable colour space. In general, the choice of colour space makes no difference to the resulting image that our network produces.

When reading images from our dataset, we first interpret them in the RBG colour space and then proceed to convert them to LAB using tools such as SciKit Learn or OpenCV, both of which contain colour space conversion libraries.

@@

When colorizing images using a deep learning approach, it is generally easiest to develop our model as a regression problem. Classification would yield better results which are more life-like and saturated whereas regression would lead to output images appearing dull and bland. This is the result of the multi-modality of pictures. There is no clear color for any one image. This leads our model, using an aggregate loss function, to simply find an average color, resulting in a bland image.

\begin{figure}[h!]
    \centering
    \subfloat[Regression]{{\includegraphics[width=3.5cm]{cham-reg} }}%
    \qquad
    \subfloat[Classification]{{\includegraphics[width=3.5cm]{cham-class} }}%
    \caption{Colour generated using regression vs classification \cite{zhang2016colorful}}%
    \label{fig:reg_vs_class}
\end{figure}

For the sake of simplicity, we continue to use regression as it leads to satisfactory results whilst being simple to understand. We utilize a mean squared error loss which may average color distribution but still produces images with a reasonable amount of color. If we aimed to produce images which were more diverse in colour and saturated, regression would not suffice. Our model should, however, be able to compare output across datasets in order to draw correlation between the accuracy of the ouput and the representivity of local images in the data, which it does.

@@

Our choice of model has been based on the simple regression model developed by Luke Melas and loosely based on the network used by Zhang et al. \cite{zhang2016colorful} with the major difference being that their model used in 'Colorful Image Colorization'\cite{zhang2016colorful} approached colorization as a classification problem.

@@

%Train colouriser on dataset X. Evaluate accuracy
%
%Explain how our dataset is built precisely: using flickr data, keywords, any preprocessing etc. 
%
%Show some images. 
%
%Maybe even compute mean statistics of dataset X and Y to show that they are in fact different?
%
%Explain method - compute accuracy on X -> X.
%Then check X -> Y
%
%Then check how much of Y must be added to X to improve things?
%
%Would be nice to have multiple datasets, models, but maybe no time :(

@@

\begin{table}[h!]
\centering
\begin{tabular}{cc}
\includegraphics[width=3.7cm]{SampleSA/46} & \includegraphics[width=3.7cm]{SampleSA/61}\\
\includegraphics[width=3.7cm]{SampleSA/67} & \includegraphics[width=3.7cm]{SampleSA/69}\\
\includegraphics[width=3.7cm]{SampleSA/106} & \includegraphics[width=3.7cm]{SampleSA/127}\\

@@

\begin{table}[h!]
\centering
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|c|c|c|c|}
\hline 
Learning Rate & Epochs & Workers & Batch Size \\ 
\hline 
0.001 & 75 & 4 & 64 \\ 
\hline 
\end{tabular} 
\end{table}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=9cm]{Curves/loss_train}
\caption{Loss curves of each dataset during training}
\label{fig:loss_train}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \subfloat[0\% inclusion]{{\includegraphics[width=3.8cm]{Inclusion_compare/0} }}%
    \qquad
    \subfloat[30\% inclusion]{{\includegraphics[width=3.8cm]{Inclusion_compare/30} }}%
    \caption{A colourised image produced using models with varying inclusion of local data}%
    \label{fig:inclusion_compare}
\end{figure}

@@

\begin{itemize}
    \item Find literature on dataset imbalance in general
    \item Find literature on racial bias in general
\end{itemize}

>>>>

\documentclass[conference]{IEEEtran}
% \IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{float}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\graphicspath{{./images/}}
\usepackage{textcomp}
\usepackage{subfig}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

@@

\subsection{Colour space}

For simlicity, rather than using the typical RGB colour space, we make use of the CIELAB colour space. This makes it convenient when seperating the gray and colour channels due to the fact that the 'L' channel in LAB is simply the grayscale image. This grayscale input (1 x W x H) is then passed through the model and compared with the output (2 x W x H) of the original 'AB' channels of the image. These 'AB' channels make up the colour and simply signify the red-green and blue-yellow shift respectively. HSB would be another option when choosing a suitable colour space. In general, the choice of colour space makes no difference to the resulting image that our network produces.

When reading images from our dataset, our chosen system interprets them in the RBG colour space and then proceeds to convert them to LAB using tools such as SciKit Learn or OpenCV, both of which contain colour space conversion libraries.

@@

When colorizing images using a deep learning approach, it is generally easiest to develop the model as a regression problem. Classification would yield better results which are more life-like and saturated whereas regression would lead to output images appearing dull and bland. This is the result of the multi-modality of pictures. There is no clear colour for any one image. This leads our model, using an aggregate loss function, to simply find an average colour, resulting in a bland image.

\begin{figure}[h]
    \centering
    \subfloat[Regression]{{\includegraphics[width=3.5cm]{cham-reg} }}%
    \qquad
    \subfloat[Classification]{{\includegraphics[width=3.5cm]{cham-class} }}%
    \caption{Colour generated using regression vs classification \cite{zhang2016colorful}}%
    \label{fig:reg_vs_class}
\end{figure}

For the sake of simplicity, we continue to use regression as it leads to satisfactory results whilst being simple to understand. We utilize a mean squared error loss which may average colour distribution but still produces images with a reasonable amount of colour. If we aimed to produce images which were more diverse in colour and saturated, regression would not suffice. Our model should, however, be able to compare output across datasets in order to draw correlation between the accuracy of the ouput and the representivity of local images in the data, which it does.

@@

Our choice of model has been based on the simple regression model developed by Luke Melas \cite{melas2018mark} and loosely based on the network used by Zhang et al. \cite{zhang2016colorful}\cite{zhang2017real} with the major difference being that their model used in 'Colorful Image Colorization'\cite{zhang2016colorful} approached colourisation as a classification problem.

@@

@@

\begin{table}[h!]
\centering
\begin{tabular}{cc}
\includegraphics[width=3.7cm]{SampleSA/46} & \includegraphics[width=3.7cm]{SampleSA/61}\\
\includegraphics[width=3.7cm]{SampleSA/106} & \includegraphics[width=3.7cm]{SampleSA/127}\\

@@

\begin{table}[h]
\centering
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|c|c|c|c|}
\hline 
Learning Rate & Epochs & Workers & Batch Size \\ 
\hline 
0.001 & 75 & 4 & 64 \\ 
\hline 
\end{tabular} 
\end{table}

@@

\begin{figure}[h]
\centering
\includegraphics[width=9cm]{Curves/loss_train}
\caption{Loss curves of each dataset during training}
\label{fig:loss_train}
\end{figure}

@@

\begin{figure}[h]
    \centering
    \subfloat[0\% inclusion]{{\includegraphics[width=3.8cm]{Inclusion_compare/0} }}%
    \qquad
    \subfloat[30\% inclusion]{{\includegraphics[width=3.8cm]{Inclusion_compare/30} }}%
    \caption{A colourised image produced using models with varying inclusion of local data}%
    \label{fig:inclusion_compare}
\end{figure}

@@

Many studies have shown that the majority of datasets used in facial analysis are comprised of mostly lighter skinned subjects. Research published by Joy Buolamwini \cite{buolamwini2018gender} shows that when evaluated on industry standard facial analysis networks, dark skinned females yield the highest error rates (of up to 34.7\%). This is alarming considering that light skinned males yield a maximum error rate of only 0.8\% when evaluated on the same dataset. \cite{corbett2018measure}

This stems from the fact that data scientists and researchers continue to neglect the unbalanced distribution of race in the datasets that they use. The systems used and sold by IBM, Microsoft and Amazon are just a few that contain these race imbalances. The systems sold by these technology giants are used by some states in the United States to apprehend criminals by means of facial recognition. The fact that there are racial imbalances in many of the networks used in these systems means that there is a higher chance for people of colour to be falsely accused of criminal activity due to facial recognition software returning a false positive.

The class imbalance problem has long been researched and there are various methods \cite{japkowicz2000class} to overcome the issue. One such method to overcome a dataset imbalance is to simply re-sample the minority until a suitable distribution is achieved. Similarly, down-sampling the majority leads to a fairer distribution as shown in \cite{japkowicz2000class}.

No substantial research has been conducted in investigating the effect of data imbalance with respect to image colourisation. This may be due to the fact that image colourisation is largely subjective as made evident in the previous section.
